---
title: "Analyse the expression data"
author: John Reid
output: pdf_document
---

```{r render, eval = FALSE, echo = FALSE}
#
# Execute this block to render this document.
#
devtools::load_all()
rmarkdown::render('expression.Rmd')
```

```{r params, echo = FALSE}
library(Saturn)
knitr::opts_chunk$set(fig.width = 18, fig.height = 12)
```


# Parameters

```{r parameters}
expr.var <- 'log.TPM'
num.expr.clusters <- 40
```


# Load data

Load expression data.
```{r loadExpr, message = FALSE}
expr <-
    data.frame(cell = factor(cell.levels, levels = cell.levels))  %>%
    group_by(cell) %>%
    do(load.avg.expr(.$cell)) %>%
    ungroup()
```

Cast as matrix.
```{r castExpr}
expr.mat <-
    expr %>%
    acast(ensembl_gene_id ~ cell, value.var = expr.var)
dim(expr.mat)
```


# Select genes

Get Ensembl IDs for genes of interest.
```{r genesInterest, message = FALSE}
library(Homo.sapiens)
ensembl.ids.of.interest <-
    select(
        Homo.sapiens,
        keys = tf.levels,
        column = 'ENSEMBL',
        keytype = 'SYMBOL',
        multiVals = 'asNA') %>%
    filter(ENSEMBL != 'ENSG00000122728')
expr.of.interest <- expr.mat[ensembl.ids.of.interest$ENSEMBL,]
```

Plot heatmap of genes of interest.
```{r heatmapInterest, message = FALSE}
library(gplots)
library(RColorBrewer)
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
hm <- heatmap.2(
    expr.of.interest,
    labRow = ensembl.ids.of.interest$SYMBOL,
    col = hmcol,
    trace = 'none')
```


# Heatmap

Pick genes with highest variances.
```{r highVar}
vars <- apply(expr.mat, 1, var)
gene.idxs <- order(-vars)[1:2000]
```

Make a heatmap of high variance genes.
```{r heatmap}
hm <- heatmap.2(
    expr.mat[gene.idxs,],
    col = hmcol,
    labRow = '',
    trace = 'none')
```


# Clustering

Extract hierarchical clustering from heatmap.
```{r hierClus}
hc.genes <- as.hclust(hm$rowDendrogram)
cluster.ids <- cutree(hc.genes, k = num.expr.clusters)
expr.clusters <- data.frame(
    ensembl_gene_id = factor(names(cluster.ids), levels = levels(expr$ensembl_gene_id)),
    cluster_id = cluster.ids)
rownames(expr.clusters) <- NULL
```

Summarise cluster expression in each cell.
```{r cellCluster}
expr.cell.cluster <-
    expr %>%
    inner_join(expr.clusters) %>%
    group_by(cell, cluster_id) %>%
    summarise(log.TPM = mean(log.TPM))
```


```{r saveData, eval = FALSE, message = FALSE, echo = FALSE}
devtools::use_data(expr.clusters, overwrite = TRUE)
devtools::use_data(expr.cell.cluster, overwrite = TRUE)
devtools::use_data(ensembl.ids.of.interest, overwrite = TRUE)
devtools::use_data(expr.of.interest, overwrite = TRUE)
```


# Session information

`r date()` on `r Sys.info()[["nodename"]]`
```{r sessionInfo}
devtools::session_info()
largest.objects()
```
