---
title: "Analyse the expression data"
author: John Reid
output: pdf_document
---

```{r render, eval = FALSE, echo = FALSE}
#
# Execute this block to render this document.
#
devtools::load_all()
rmarkdown::render('features-expr.Rmd')
```

```{r params, echo = FALSE}
library(Saturn)
# library(edgeR)
library(stringr)
library(ggplot2)
library(ggthemes)
library(GenomicRanges)
library(S4Vectors)
knitr::opts_chunk$set(fig.width = 18, fig.height = 12)
```


```{r writeBed, eval = FALSE, echo = FALSE}
rtracklayer::export(ranges.test(), '../Data/annotations/test_regions.ordered.bed', 'BED')
```


# Load data

Load expression data into data.frame.
```{r loadExpr, message = FALSE}
load_expr <- function(cell) {
    .df <-
        rbind(
            saturn.expr(cell, 1) %>%
                dplyr::select(gene = gene_id, TPM) %>%
                dplyr::mutate(rep = 1, cell = cell),
            saturn.expr(cell, 2) %>%
                dplyr::select(gene = gene_id, TPM) %>%
                dplyr::mutate(rep = 2, cell = cell)) %>%
        dcast(gene ~ rep, value.var = 'TPM') %>%
        setNames(c('gene', 'TPM1', 'TPM2')) %>%
        mutate(cell = cell, TPM.avg = (TPM1 + TPM2) / 2)
}
expr <- do.call(rbind, lapply(cells$cell, load_expr))
```

Extract `ENSG` identifiers
```{r extractENSG}
expr$ENSGENE <- str_split_fixed(expr$gene, fixed('.'), 2)[,1]
```

# Compare replicates

```{r compareReps}
dim(expr)
ggplot(expr, aes(x = TPM1, y = TPM2)) +
    geom_point(alpha = .1) +
    facet_wrap(~ cell) +
    scale_x_log10() +
    scale_y_log10() +
    theme_few()
ggsave('../Plots/expr-by-rep.png')
```


# Compare cells

How many zeros are there in each cell?
```{r cellZeros}
cell.zeros <-
    expr %>%
    group_by(cell) %>%
    summarise(N = n(), Nz = sum(0 == TPM.avg))
ggplot(cell.zeros, aes(x = cell, y = Nz)) +
    geom_bar(stat = 'identity') +
    theme_few()
```

Compare expression densities across cells
```{r compareCells}
ggplot(expr %>% filter(TPM.avg > 0), aes(x = TPM.avg)) +
    geom_density(fill = 'black') +
    scale_x_log10() +
    facet_wrap(~ cell) +
    theme_few()
ggsave('../Plots/expr-by-cell.pdf')
```


# Distance to closest gene

Load closest gene information
```{r closestGene}
nearest <- rtracklayer::import('../Data/annotations/closest-genes.bed.gz')
start(nearest) <- start(nearest) - 1
end(nearest) <- end(nearest) - 1
seqlevels(nearest) <- seqlevels(ranges.test())
seqinfo(nearest) <- seqinfo(ranges.test())
nearest <- sort(nearest)
names(mcols(nearest)) <- c('ENSGENE', 'distance')
mcols(nearest)$ENSGENE <- Rle(mcols(nearest)$ENSGENE)
```

Examine distribution of the distance to the nearest gene:
```{r examDist}
qplot(mcols(sample(nearest, 100000))$distance) + scale_x_log10() + theme_few()
```

Choose a length scale of `1e4`:
```{r lengthScale}
lengthscale <- 1e4
mcols(nearest)$similarity <- exp(-(mcols(nearest)$distance/lengthscale)**2/2)
qplot(mcols(sample(nearest, 10000))$similarity) + theme_few()
```

Bin the distances and the similarities:
```{r binDists}
n.bins <- 20     # How many distinct values will our features have?
epsilon <- 1e-8  # Small adjustment to widen range
nearest.feat <-
    with(mcols(nearest),
        DataFrame(
            similarity = Rle(as.numeric(cut(similarity, seq(-epsilon, 1+epsilon, length.out = n.bins + 1)))),
            distance = Rle(as.numeric(cut(log10(distance+1), seq(-epsilon, log10(max(distance+1))+epsilon, length.out = n.bins + 1))))))
sum(is.na(nearest.feat$similarity))
sum(is.na(nearest.feat$distance))
```

Save distance features:
```{r saveDist}
saveRDS(nearest.feat, '../Data/Features/GeneDist/GeneDist.rds')
```


# Expression of closest gene

```{r nearestExpr}
for (.cell in cells$cell) {
    message(.cell)
    cell.expr <-
        as.data.frame(mcols(nearest)) %>%
        left_join(
            expr %>% filter(cell == .cell) %>% dplyr::select(ENSGENE, TPM.avg))
    saveRDS(
        DataFrame(nearest.expr = Rle(cell.expr$TPM.avg)),
        stringr::str_c('../Data/Features/NearExpr/NearExpr-', .cell, '.rds'))
}
```


# Session information

`r date()` on `r Sys.info()[["nodename"]]`
```{r sessionInfo}
devtools::session_info()
largest.objects()
```
