---
title: "Get negative and position FASTA sequences for a TF"
author: John Reid
output: html_document
params:
  TF: CTCF
---

```{r renderMe, eval = FALSE, echo = FALSE}
#
# Execute this block to render this document.
#
devtools::load_all()
if (FALSE) {
    rmarkdown::render("get-TF-fasta.Rmd", params = list(TF = "CTCF"))
} else {
    for (TF in levels(tfs$TF)) {
        rmarkdown::render(
            "get-TF-fasta.Rmd",
            output_file = str_c('FASTA-', TF, '.html'),
            params = list(TF = TF))
    }
}
```

Load the ChIP data for `{r params$TF}`.
```{r loadChip}
library(Saturn)
TF <- params$TF
chip <- read.chip.labels(TF)
```

Analyse the ChIP data to see which regions bind in at least one cell type.
```{r getPositive}
tf.cells <- cell.names(chip)
is.bound <- Map(function(x) 'B' == x, as.list(chip[tf.cells]))
bound <- Reduce(function(x,y) x | y, is.bound)
# Check we have at least as many unbound sequences
stopifnot(sum(! bound) >= sum(bound))
# Generate ranges
positive <- regions.to.ranges(chip[bound,])
# Make sure we have right number
stopifnot(sum(bound) == length(positive))
```
We have `r length(positive)` positive ChIP regions.

Determine maximum DNase levels across cell types.
```{r dnase}
dnase.cells <- lapply(as.character(tf.cells), function(cell) summarise.dnase(cell, 'relaxed'))
dnase.max <- Reduce(pmax, dnase.cells)
```

Sort the training regions by their maximum DNase levels and select those with
the highest levels that are not bound
```{r getNegative}
# Get the maximum DNase levels in the training regions
dnase.max.idx <- order(dnase.max[training.region.test.idxs()], decreasing = TRUE)
# Limit to the regions that are unbound and only as many as in the positive set
unbound.dnase.max.idx <- dnase.max.idx[as.vector(! bound)][1:length(positive)]
# Find the negative regions with the lowest and highest dnase levels
highest.neg.dnase <- dnase.max[train.to.test.idx(unbound.dnase.max.idx[1])]
lowest.neg.dnase  <- dnase.max[train.to.test.idx(unbound.dnase.max.idx[length(positive)])]
message('First negative region has DNase level: ', highest.neg.dnase)
message('Last  negative region has DNase level: ', lowest.neg.dnase)
# Generate ranges
negative <- regions.to.ranges(chip[unbound.dnase.max.idx,])
# Check we have correct number
stopifnot(length(negative) == length(positive))
```
We have `r length(negative)` negative ChIP regions.

Write the positive and negative sequences to separate FASTA files.
```{r writeSeqs}
write.seqs <- function(gr, label) {
    seqs <- BSgenome::getSeq(hg19, gr)
    names(seqs) <- stringr::str_c(TF, ".", label, ".", rtracklayer::chrom(gr), ':', start(gr), ':', end(gr))
    fasta <- file.path(saturn.data(), 'ChIPseq', 'seqs', stringr::str_c(TF, '-', label, '.fa'))
    Biostrings::writeXStringSet(seqs, fasta)
}
write.seqs(positive, 'pos')
write.seqs(negative, 'neg')
```
