---
title: "Ladderboard, training and test regions"
author: John Reid
output: html_document
---

```{r renderMe, eval = FALSE, echo = FALSE}
#
# Execute this block to render this document.
#
devtools::load_all()
rmarkdown::render("regions.Rmd")
```


# Load

Load the regions from the annotation directory.
```{r load}
library(Saturn)
#
# Ladder regions
regions.ladder <- load.regions('ladder')
object.size(regions.ladder)
#
# Train regions
regions.train <- load.regions('train')
object.size(regions.train)
#
# Test regions
regions.test <- load.regions('test')
object.size(regions.test)
```
Have `r length(regions.ladder)` ladder regions, `r length(regions.train)` train regions and `r length(regions.test)`
test regions.

Double-check that the test regions are the union of the ladderboard and test regions.
```{r}
.union <- GenomicRanges::union(ranges.ladder(), ranges.train())
stopifnot(length(GenomicRanges::setdiff(.union, ranges.test())) == 0)
```

Double-check that the ladderboard and train regions don't intersect.
```{r}
stopifnot(length(GenomicRanges::intersect(ranges.ladder(), ranges.train())) == 0)
```

Calculate the indices of the training and ladder regions in the test regions
```{r trainIdxs}
regions.as.dt <- function(regions) with(regions,
  setkey(
    data.table(
      chrom = as.factor(chrom),
      start = as.integer(start)),
    chrom,
    start))
ladder <- regions.as.dt(regions.ladder) %>% mutate(ladder.idx = 1:nrow(regions.ladder))
train  <- regions.as.dt(regions.train)  %>% mutate(train.idx  = 1:nrow(regions.train))
test   <- regions.as.dt(regions.test)   %>% mutate(test.idx   = 1:nrow(regions.test))
regions.train$test.idx  <- merge(test, train )$test.idx
regions.ladder$test.idx <- merge(test, ladder)$test.idx
regions.test$ladder.idx <- merge(test, ladder, all.x = TRUE)$ladder.idx
regions.test$train.idx  <- merge(test, train , all.x = TRUE)$train.idx
```


# Save

Save the regions
```{r save, eval = FALSE}
devtools::use_data(regions.ladder, overwrite = TRUE)
devtools::use_data(regions.train, overwrite = TRUE)
devtools::use_data(regions.test, overwrite = TRUE)
```

# Session information

`r date()`
```{r sessionInfo}
devtools::session_info()
```
