---
title: "ENCODE DREAM Challenge"
author: John Reid
output: pdf_document
---

```{r echo=FALSE, eval=FALSE}
rmarkdown::render('ENCODE-DREAM-overview.Rmd')
```


```{r parseCombos, echo=FALSE}
library(stringr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggthemes)
#
# Parse the information about the TFs.
scraped <-
'ARID3A 	HepG2 	K562 	
ATF2 	GM12878, H1-hESC, MCF-7 	K562 	HepG2
ATF3 	HCT116, H1-hESC, HepG2, K562 	liver 	
ATF7 	GM12878, HepG2, K562 	MCF-7 	
CEBPB 	A549, H1-hESC, HCT116, HeLa-S3, HepG2, IMR-90, K562 	MCF-7 	
CREB1 	GM12878, H1-hESC, HepG2, K562 	MCF-7 	
CTCF 	A549, H1-hESC, HeLa-S3, HepG2, IMR-90, K562, MCF-7 	GM12878 	PC-3, induced_pluripotent_stem_cell
E2F1 	GM12878, HeLa-S3 		K562
E2F6 	A549, H1-hESC, HeLa-S3 	K562 	
EGR1 	GM12878, H1-hESC, HCT116, MCF-7 	K562 	liver
EP300 	GM12878, H1-hESC, HeLa-S3, HepG2, K562, SK-N-SH 	MCF-7 	
FOXA1 	HepG2 	MCF-7 	liver
FOXA2 	HepG2 		liver
GABPA 	GM12878, H1-hESC, HeLa-S3, HepG2, MCF-7, SK-N-SH 	K562 	liver
GATA3 	A549, SK-N-SH 	MCF-7 	
HNF4A 	HepG2 		liver
JUND 	HCT116, HeLa-S3, HepG2, K562, MCF-7, SK-N-SH 	H1-hESC 	liver
MAFK 	GM12878, H1-hESC, HeLa-S3, HepG2, IMR-90 	K562, MCF-7 	
MAX 	A549, GM12878, H1-hESC, HCT116, HeLa-S3, HepG2, K562, SK-N-SH 	MCF-7 	liver
MYC 	A549, HeLa-S3, K562, MCF-7 	HepG2 	
NANOG 	H1-hESC 		induced_pluripotent_stem_cell
REST 	H1-hESC, HeLa-S3, HepG2, MCF-7, Panc1, SK-N-SH 	K562 	liver
RFX5 	GM12878, HeLa-S3, MCF-7, SK-N-SH 	HepG2 	
SPI1 	GM12878 	K562 	
SRF 	GM12878, H1-hESC, HCT116, HepG2, K562 	MCF-7 	
STAT3 	HeLa-S3 	GM12878 	
TAF1 	GM12878, H1-hESC, HeLa-S3, K562, SK-N-SH 	HepG2 	liver
TCF12 	GM12878, H1-hESC, MCF-7, SK-N-SH 	K562 	
TCF7L2 	HCT116, HeLa-S3, Panc1 	MCF-7 	
TEAD4 	A549, H1-hESC, HCT116, HepG2, K562, SK-N-SH 	MCF-7 	
YY1 	GM12878, H1-hESC, HCT116, HepG2, SK-N-SH 	K562 	
ZNF143 	GM12878, H1-hESC, HeLa-S3, HepG2 	K562 	'
lines <- str_split(scraped, fixed('\n'))[[1]]
parse.line <- function(l) {
    # print(l)
    parsed <- str_trim(str_split(l, fixed('\t'), 4)[[1]])
    # parsed[[2]] <- str_split(parsed[[2]], fixed(', '))[[1]]
    # print(parsed)
    parsed
}
parsed <- lapply(lines, parse.line)
type.for.col <- function(col) {
    if (2 == col) return('train')
    if (3 == col) return('ladder')
    if (4 == col) return('submit')
}
parse.tfs <- function(parsed, col) {
    data.frame(
        TF=parsed[[1]],
        cell=str_split(parsed[[col]], fixed(', '))[[1]],
        type=type.for.col(col))
}
parse.cols <- function(parsed) {
    res <- parse.tfs(parsed, 2)
    if (length(parsed) > 2) res <- rbind(res, parse.tfs(parsed, 3))
    if (length(parsed) > 3) res <- rbind(res, parse.tfs(parsed, 4))
    res %>% filter("" != cell)
}
tfs <- do.call(rbind, lapply(parsed, parse.cols))
```


# The ENCODE-DREAM challenge

The [challenge](https://www.synapse.org/#!Synapse:syn6131484/wiki/) is to
predict cell-type specific binding of transcription factors (TFs) using four
types of data:

- DNA sequence: a reference human genomic sequence
- *In-vitro* DNA shape: the physical shape of the genome in an *in-vitro* system
- DNase-seq: how open and accessible regions of the genome are
- RNA-seq: levels of gene expression


## TF/cell-type combinations

TFs have different binding profiles in different cell types. `r length(unique(tfs$TF))`
TFs and `r length(unique(tfs$cell))` cell types are represented in the data although not
all combinations are present.
```{r, echo=FALSE}
ggplot(tfs, aes(TF, cell)) + geom_tile(aes(fill=type), colour = "white") +
    scale_fill_few() +
    theme_tufte() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
There are essentially three (two!?) prediction tasks:
- Held-out chromosomes: For each TF/cell-type combination, the ChIP-seq data will
  not be available for 3 chromosomes. Methods will be assessed by their predictive
  performance on these chromosomes.
- Across cell-type: Methods will be assessed by their performance on cell-types for
  which no training data has been made available (PC-3, induced pluripotent stem
  cells, liver).
- Within cell-type: Methods will be assessed on 


## Data


### TF binding

TF binding is measured using the ChIP-seq protocol and converted to a binary
score for sliding windows of 200bp. The windows slide by 50bp. For each 200bp
location, binding is defined as bound (B), unbound (U) or ambiguous (A). In
addition the challenge provides more detailed information from the ChIP-seq
experiments including conservative and relaxed estimates of peaks and
fold-control signals showing how enriched the ChIP-seq experiment was over a
control background experiment. It is not clear how useful the extra information
will be as it will obviously not be available on the held-out data.


### DNA sequence

The human reference genome is over 3 billion base pairs long. Each base is
represented as a character: adenine (A); cytosine (C); guanine (G) and thymine
(T). The challenge is restricted to chromosomes 1-22 and chromosome X. All data
in the challenge are defined with respect to release GRCh37/hg19.  TFs tend to
prefer to bind specific sequences. These preferences are summarised in binding
motifs but these are not known for all TFs. The only external data that is
allowed to be used in this challenge are libraries of TF-DNA binding motifs.
Obviously this data is not cell-type specific.


### DNase-seq

Information on chromatin accessibility on a per-cell-type basis will be
available in four formats:

- conservative peaks
- relaxed peaks
- filtered BAM alignment files
- fold-enrichment signal coverage tracks

The first two will be much easier to use as they summarise the last two.


### DNA shape

Participants are encouraged to use the [DNAshapeR](http://tsupeichiu.github.io/DNAshapeR/) to
calculate DNA shape features across the genome. This information is not cell-type specific but
has been shown to be predictive of TF-DNA binding.


### Gene expression

Gene expression is regulated by TF binding and so should be useful indirectly to infer binding.
The major difficulties are that
- It is known which locations on the genome regulate which genes. Commonly TF binding sites
  regulate the closest gene but this is not always the case.
- Several TFs can combine in an unknown way to regulate a gene.
- There are other mechanisms of gene regulation that will confound the relationship.

